\section{Templates for Schur numbers}
\label{Schur}

In this section, we use Rowley's template-based constructions \cite{RowleyRamsey} in the context of Schur 
numbers. In order to improve lower bounds for Schur and Ramsey numbers, he introduces special sum-free 
partitions verifying some additional properties which can be extended using a method generalizing Abbott and 
Hanson's construction \cite{AbbottHanson}. He named these partitions "templates", and we keep this name in 
the entire article. We then find new templates and use them to provide new lower bounds for Schur numbers.

\subsection{Definition of \( [\![ 1, S^+ ]\!] \)}

We begin by introducing \textit{S-templates}, standing for \textit{"Schur templates"}. The idea is to consider the first line
of \hyperref[figure:1]{Figure 1} not as a combination of two blocs anymore but as a whole, single construction. A S-template 
is then defined as a new object filling the role of the first line but with less, yet sufficiant, constraints which allow a linear 
expansion of the partition.

\begin{definition}
Let \((p,n) \in (\mathbb{N}^*)^2\). A S-template with width \(p\) and \(n\) colors is defined as a partition of 
\([\![1,p]\!]\) into \(n\) sum-free subsets \(A_1, A_2, ..., A_n\) verifying for all subsets but \(A_n\)
\[
\forall i \in [\![1, n-1]\!], \forall (x,y) \in A_i^2, x+y > p
\Longrightarrow x+y-p \notin A_i
\]
\end{definition}

Here \(n\) can be seen as a "special" color in the sense that we do not impose this above additionnal constraint on this color. 
However, note that \(n\) is not necessarily the last color by order of appearance, any color can play this role.


\begin{proposition}
	Let \(n \in [\![2, +\infty[\![\). We define \(S^+(n)\) as the greatest integer such that \( [\![1, S^+(n)]\!]\) is a S-template with \(n\) colors. 
	\(S^+(n)\) is well defined and verifies
	\[
	2S(n-1)+1 \leqslant S^+(n) \leqslant S(n)
	\]
\end{proposition}

\begin{proof}[\textsc{Proof.}]
The lower bound comes from Abbott and Hanson's construction. The upper bound comes from the 
fact that a S-template with width \(p\) and \(n\) colors is also a partition of \([\![1, p]\!]\) into \(n\) sum-free subsets. \\
\end{proof}


\subsection{Construction of Schur partitions using S-templates}

We start by reminding the explicit construction of a sum-free partition with the use of a S-template and another sum free partition. We apply this contruction stated by Rowley in the context of Ramsey numbers to the Schur numbers.

\begin{theorem}
\label{thm:Stemp}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there exists a \(S\)-template with width \(q\) and \(n+1\) colors 
	and a partition of \([\![1,p]\!]\) into \(k\) sum-free subsets then there exists a partition of \([\![1,pq+m_{n+1}
	-1]\!]\) into \(n+k\) sum-free subsets. \(m_{n+1}\) denotes the minimum element colored with the special color in the \(S\)-template.
\end{theorem}

Setting \(q = S^+(n+1)\) and \(p = S(k)\) yields the following corollary.

\begin{corollary}
\label{cor:ineqS}
	Let \(n, k \in \mathbb{N}^*\). Then
	\[ S(n+k) \geqslant S^+(n+1)S(k) + m_{n+1} - 1 \]
\end{corollary}

The idea lying beneath this theorem is similar to Abbott and Hanson's contruction \cite{AbbottHanson}. They vertically 
extend a sum-free partition by repeating it and they use an other sum-free partition to color the other half according 
to the line number. This way the "blocks" act as safe areas for each other. We give here an example for \(p = 4\), 
\(q = 9\), \(n = 2\) and \(k = 2\) showing that \(S(2 + 2) \geqslant S(2) (2 S(2) + 1) + S(2)\), both with Abbott and 
Hanson's construction and with a S-template which is not included in Abbott and Hanson's construction. In both cases, 
the special color is grey.

\renewcommand{\arraystretch}{1.7}
\begin{figure}[H]
\begin{center}
\setlength{\tabcolsep}{1.5ex}

\textbf{Abbott and Hanson's construction}
\vspace{1.7ex}

\begin{NiceTabular}{|*{9}{c|}}[corners=SE,standard-cline,hlines]
\CodeBefore
	\cellcolor{red}{1-1}
	\cellcolor{green}{1-2}
	\cellcolor{green}{1-3}
	\cellcolor{red}{1-4}
	\cellcolor{cyan}{1-5}
	\cellcolor{cyan}{1-6}
	\cellcolor{cyan}{1-7}
	\cellcolor{cyan}{1-8}
	\cellcolor{cyan}{1-9}
	\cellcolor{red}{2-1}
	\cellcolor{green}{2-2}
	\cellcolor{green}{2-3}
	\cellcolor{red}{2-4}
	\cellcolor{yellow}{2-5}
	\cellcolor{yellow}{2-6}
	\cellcolor{yellow}{2-7}
	\cellcolor{yellow}{2-8}
	\cellcolor{yellow}{2-9}
	\cellcolor{red}{3-1}
	\cellcolor{green}{3-2}
	\cellcolor{green}{3-3}
	\cellcolor{red}{3-4}
	\cellcolor{yellow}{3-5}
	\cellcolor{yellow}{3-6}
	\cellcolor{yellow}{3-7}
	\cellcolor{yellow}{3-8}
	\cellcolor{yellow}{3-9}
	\cellcolor{red}{4-1}
	\cellcolor{green}{4-2}
	\cellcolor{green}{4-3}
	\cellcolor{red}{4-4}
	\cellcolor{cyan}{4-5}
	\cellcolor{cyan}{4-6}
	\cellcolor{cyan}{4-7}
	\cellcolor{cyan}{4-8}
	\cellcolor{cyan}{4-9}
	\cellcolor{red}{5-1}
	\cellcolor{green}{5-2}
	\cellcolor{green}{5-3}
	\cellcolor{red}{5-4}
\Body
	1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
	10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 \\
	19 & 20 & 21 & 22 & 23 & 24 & 25 & 26 & 27 \\
	28 & 29 & 30 & 31 & 32 & 33 & 34 & 35 & 36 \\
	37 & 38 & 39 & 40 \\
\end{NiceTabular}

\vspace{1ex}
\setlength{\tabcolsep}{2ex}

\begin{tabular}{c c}
	\textbf{Corresponding S-template} & \textbf{Corresponding sum-free partition} \\
	\begin{NiceTabular}{|*{9}{c|}}[standard-cline,hlines]
	\CodeBefore 
		\cellcolor{red}{1-1}
		\cellcolor{green}{1-2}
		\cellcolor{green}{1-3}
		\cellcolor{red}{1-4}
		\cellcolor{gray!40}{1-5}
		\cellcolor{gray!40}{1-6}
		\cellcolor{gray!40}{1-7}
		\cellcolor{gray!40}{1-8}
		\cellcolor{gray!40}{1-9}
	\Body
		1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
	\end{NiceTabular} &
	\begin{NiceTabular}{|*{4}{c|}}[standard-cline,hlines]
	\CodeBefore
		\cellcolor{cyan}{1-1}
		\cellcolor{yellow}{1-2}
		\cellcolor{yellow}{1-3}
		\cellcolor{cyan}{1-4}
	\Body
		1 & 2 & 3 & 4 \\
	\end{NiceTabular}
\end{tabular}

\setlength{\tabcolsep}{6pt}
\caption{Visualisation of an Abbott and Hanson construction}
\end{center}
\label{figure:1}
\end{figure}
In the general construction with S-templates, the special color no longer necessarily contains consecutive 
numbers. However, the special color is still replaced by the colors of the sum-free partition according to the 
line number and the other colors are still vertically extended.

\begin{figure}[H]
\begin{center}
\textbf{S-template construction}
\setlength{\tabcolsep}{1.5ex}

\vspace{1.7ex}

\begin{NiceTabular}{|*{9}{c|}}[corners=SE,standard-cline,hlines]
\CodeBefore
	\cellcolor{red}{1-1}
	\cellcolor{green}{1-2}
	\cellcolor{green}{1-3}
	\cellcolor{red}{1-4}
	\cellcolor{cyan}{1-5}
	\cellcolor{cyan}{1-6}
	\cellcolor{red}{1-7}
	\cellcolor{cyan}{1-8}
	\cellcolor{cyan}{1-9}
	\cellcolor{red}{2-1}
	\cellcolor{green}{2-2}
	\cellcolor{green}{2-3}
	\cellcolor{red}{2-4}
	\cellcolor{yellow}{2-5}
	\cellcolor{yellow}{2-6}
	\cellcolor{red}{2-7}
	\cellcolor{yellow}{2-8}
	\cellcolor{yellow}{2-9}
	\cellcolor{red}{3-1}
	\cellcolor{green}{3-2}
	\cellcolor{green}{3-3}
	\cellcolor{red}{3-4}
	\cellcolor{yellow}{3-5}
	\cellcolor{yellow}{3-6}
	\cellcolor{red}{3-7}
	\cellcolor{yellow}{3-8}
	\cellcolor{yellow}{3-9}
	\cellcolor{red}{4-1}
	\cellcolor{green}{4-2}
	\cellcolor{green}{4-3}
	\cellcolor{red}{4-4}
	\cellcolor{cyan}{4-5}
	\cellcolor{cyan}{4-6}
	\cellcolor{red}{4-7}
	\cellcolor{cyan}{4-8}
	\cellcolor{cyan}{4-9}
	\cellcolor{red}{5-1}
	\cellcolor{green}{5-2}
	\cellcolor{green}{5-3}
	\cellcolor{red}{5-4}
\Body
	1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
	10 & 11 & 12 & 13 & 14 & 15 & 16 & 17 & 18 \\
	19 & 20 & 21 & 22 & 23 & 24 & 25 & 26 & 27 \\
	28 & 29 & 30 & 31 & 32 & 33 & 34 & 35 & 36 \\
	37 & 38 & 39 & 40 \\
\end{NiceTabular}

\vspace{1ex}
\setlength{\tabcolsep}{2ex}

\begin{tabular}{c c}
	\textbf{Corresponding S-template} & \textbf{Corresponding sum-free partition} \\
	\begin{NiceTabular}{|*{9}{c|}}[standard-cline,hlines]
	\CodeBefore 
		\cellcolor{red}{1-1}
		\cellcolor{green}{1-2}
		\cellcolor{green}{1-3}
		\cellcolor{red}{1-4}
		\cellcolor{gray!40}{1-5}
		\cellcolor{gray!40}{1-6}
		\cellcolor{red}{1-7}
		\cellcolor{gray!40}{1-8}
		\cellcolor{gray!40}{1-9}
	\Body
		1 & 2 & 3 & 4 & 5 & 6 & 7 & 8 & 9 \\
	\end{NiceTabular} &
	\begin{NiceTabular}{|*{4}{c|}}[standard-cline,hlines]
	\CodeBefore
		\cellcolor{cyan}{1-1}
		\cellcolor{yellow}{1-2}
		\cellcolor{yellow}{1-3}
		\cellcolor{cyan}{1-4}
	\Body
		1 & 2 & 3 & 4 \\
	\end{NiceTabular}
\end{tabular}

\setlength{\tabcolsep}{6pt}
\caption{Visualisation of a template-based construction}
\label{figure:2}

\end{center}
\end{figure}

\resetarraystretch

We now proceed to prove Theorem \ref{thm:Stemp}.

\begin{proof}[\textsc{Proof.}]
	Denote by \(f\) the coloring associated to the \(S\)-template with width \(q\) and \(g\) the one associated 
	to the sum-free partition of \([\![1,p]\!]\); where \(f : [\![1,q]\!] \longrightarrow [\![1,n+1]\!]\) and 
	\(g : [\![1,p]\!] \longrightarrow [\![1,k]\!]\).
	
	\par
	NB: In the following three predicates, the conditions \(x + y \leqslant q\)  and \(x + y \leqslant p\) are omitted for readability.
	\par
	The sum-free condition is expressed as:
	\[\forall (x,y) \in [\![1,q]\!]^2, f(x) = f(y) \Longrightarrow f(x+y) \neq f(x),\]
	\[\forall (x,y) \in [\![1,p]\!]^2, g(x) = g(y) \Longrightarrow g(x+y) \neq g(x).\]
	
	The additionnal constraint for the S-template is:
	\[
	\forall (x,y) \in [\![1,q]\!]^2, \left\{
	\begin{array}{l}
		f(x) = f(y) \leqslant n \\
		x + y > q
	\end{array}
	\right. \Longrightarrow f(x+y-q) \neq f(x).
	\]
	
	For \(x \in [\![1,pq+m_{n+1}-1]\!]\), we write \(x = (\alpha - 1) + u\) for certain integers \(\alpha \in \mathbb{Z}\) and 
	\(u \in [\![1,q]\!]\). This decomposition is of course unique. \(\alpha\)  can be interpreted as the row number of \(x\) and 
	\(u\) as the collumn number of \(x\). A new coloring is define as follows:
	\[
	\begin{array}{c c c l}
		h : & [\![1,pq+m_{n+1}-1]\!] & \longrightarrow & [\![1,n+k]\!], \\
		& x & \longmapsto & 
		\left\{ \begin{array}{l l}
			f(u), & \text{if}~f(u) \leqslant n, \\
			n + g(\alpha), & \text{if}~f(u) = n + 1. \\
		\end{array} \right.
	\end{array}
	\]
	
	Function \(h\) is well-defined since, by definition of \(m_{n+1}\), \(\forall x \in [\![p q + 1, p q + m_{n + 1} - 1 ]\!], f(u) 
	\leqslant n\) and therefore \(\forall x \in [\![1,pq+m_{n+1}-1]\!], f(u) = n + 1 \implies \alpha \in [\![1, p]\!]\).
	
	\par
	We now prove that \(h\) is a sum-free coloring. Let \(x,y \in [\![1,pq + m_{n+1}-1]\!]\) such that \(h(x) = h(y)\) 
	and \(x+y \leqslant pq+m_{n+1}-1\). We claim that \(h(x+y) \neq h(x)\). We write \(x = (\alpha - 1) q + u\) and 
	\(y = (\beta - 1) q + v\) where \(\alpha, \beta \in \mathbb{Z}\) and \(u, v \in [\![1,q]\!]\). Two cases are to be
	distinguished according to the value of \(h(x)\). \\
	
	\noindent \underline{\textbf{Case 1:} \(h(x) \leqslant n\)}
	\par
	Let us assume that \(h(x+y) \leqslant n\), otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition of function 
	\(h\) and given that \(h(u) = h(v)\), we conclude \(f(u) = f(v)\). Two cases are to b distinguished according to the value of \(x + y\).
	
	\begin{itemize}
	\item If \(u + v > q\), we write \(w = u + v - q \in [\![1, q]\!]\). Consequentely \(x + y = (\alpha + \beta - 1) q + w\). By definition, 
		\(h(x + y) = f(w)\). Given that \(f(u) = f(v) \leqslant n\), the additionnal constraint on \(f\) implies \(f(w) 
		\neq f(u)\), that is \(h(x + y) \neq h(x)\).
	\item If \(u + v \leqslant p\), we write \(w = u + v \in [\![1, q]\!]\). Consequentely \(x+y = (\alpha + \beta- 2) q + w\). By definition, 
		\(h(x + y) = f(w)\). Given that \(f(u) = f(v) \leqslant n\), the sum-free property of \(f\) implies \(f(w) \neq f(u)\), 
		that is \(h(x + y) \neq h(x)\).
	\end{itemize} 
	  
	\noindent \underline{\textbf{Case 2:} \(h(x) \geqslant n + 1\)}
	\par
	Now we have \(h(x) = n + g(\alpha) = 
	k + g(\beta) = h(y)\), hence \(g(\alpha) = g(\beta)\). As in the first case, distinguish between two cases according 
	to the value of \(x + y\).
	
	\begin{itemize}
	\item \begin{sloppypar}
		If \(u + v > q\), write \(w = u + v - q \in [\![1, q]\!]\). Then \(x + y = (\alpha + \beta - 1) q + w\). Assume that 
		\({h(x+y) \geqslant n + 1}\),  otherwise \(h(x + y) \neq h(x)\) obviously holds. By definition, \({h(x + y) = n + g(\alpha + 
		\beta)}\). Given that \(g(\alpha) = g(\beta)\), the sum-free property of \(g\) implies \(g(\alpha + \beta) 
		\neq g(\alpha)\) that is \(h(x + y) \neq h(x)\).
		\end{sloppypar}
	\item  If \(u + v \leqslant q\), write \(w = u + v \in [\![1, q]\!]\). Then \(x+y = (\alpha + \beta- 2) q + w\). The sum-free 
		property of \(f\) implies \(f(w) \neq f(u)\). Therefore \(f(w) \leqslant k\) and thus \(h(x + y) \leqslant n\). In particular,
		given that \(h(x) \geqslant n + 1\), \(h(x + y) \neq h(x)\).
	\end{itemize}
\end{proof}

The following proposition may improve the additive constant of Corollary \ref{cor:ineqS}.

\begin{proposition}
\label{prop:rafStemp}
Let \((q, n) \in \mathbb{N}^*)^2\) and let \(f\) be a coloring associated to a S-template with width \(q\) and \(n+1\) 
colors. Let \(b \in \mathbb{N}\) and assume there is a coloring \(g\) of 
\([\![1, b]\!]\) with \(n+1\) colors such that:
	
\begin{itemize}
\item \(\forall (x, y) \in [\![1, q]\!]^2, \left\{
	\begin{array}{l}
		f(x) = f(y) \\
		(x + y) \mod q \leqslant b
	\end{array}
	\right. \implies g((x + y) \mod q) \neq f(x)\),
\item \(\forall (x, y) \in [\![1, q]\!] \times  [\![1, b]\!],  \left\{
	\begin{array}{l}
		f(x) = g(y) \\
		x + y \leqslant b
	\end{array}
	\right. \implies g(x + y) \neq f(x)\).
\end{itemize}
	
Then, for every \(n \in \mathbb{N}^*\), by using on the last row the coloring \(x \longmapsto g(x - p S(n))\), we have
\[ S(n+k) \geqslant S^+(n+1)S(k) + b\].
\end{proposition}
This proposition corresponds to the fact that the hypotheses made on the coloring of the last row can be weakened and as a result, we can change the coloring on the last row to extend the previous partitions.
\newline

There is a construction theorem for S-templates as well.

\begin{theorem}
	Let \((p,k), (q,n) \in (\mathbb{N}^*)^2\). If there is a S-template with width \(q\) and \(n+1\) colors,
	and a S-template with width \(p\) and \(k\) colors, then there also is a S-template with width \(pq\) and \((n+k)\) 
	colors.
	\end{theorem}
	
	The inequality associated with this theorem is given by:
	
\begin{corollary}
	Let \(n, k \in \mathbb{N}^*\). Then
	\[ S^+(n+k) \geqslant S^+(n+1)S^+(k) \].
\end{corollary}

\begin{proof}[\textsc{Proof.}]
The idea is the same as in the Theorem \ref{thm:Stemp}. The only difference is the S-template property inherited 
from the second S-template. \\
\end{proof}


\subsection{New lower bounds for Schur numbers}
\label{subsec:lowS}

We now give the inequalities corresponding to the current  best S-templates.

\begin{definition}
A sum-free partition \(A_1, ..., A_n\) of \([\![1, p]\!]\) is said to be symmetric if for all \( x \in [\![1, p]\!]\), 
\(x\) and \(p + 1 - x\) belong to the same subset (except if \(x = p + 1 - x\)).

A S-template with \(n\) colors is said to be symmetric if the partition into \(n\) sum-free subsets derived 
from this template by applying the extension procedure with a sum-free partition of length 1 is symmetric. 
\end{definition}

We produced S-templates using a SAT solver, hence providing lower bound on \(S^+\) and inequalities 
of the type \(S(n+k) \geqslant a S(n) + b\). We sought templates providing the largest value of 
\((a, b)\) (in the lexicographic order). When the number of colors exceeded five, in order to reduce the search space we 
only looked for symmetric S-templates, we assumed that the special color was the last color to appear and we constrained 
the \(m_c\)'s out of being too small. Details concerning the encoding as a SAT problem can be found in \cite{Heule2017}.

\par
Here are the inequalities given by the current best S-templates. The \hyperref[S-templates]{templates} 
corresponding to the third, fourth and fifth inequalities can be found in the appendix.

\begin{align}
	S(n + 1) \geqslant &~3\,S(n)  + 1 \label{S(n+1)}\\
	S(n + 2) \geqslant &~9\,S(n)  + 4 \label{S(n+2)}\\
	S(n + 3) \geqslant &~33\,S(n) + 6 \label{S(n+3)}\\
\end{align}

Inequality (\ref{S(n+1)}) comes from  Schur's original article\cite{Schur1917}. Inequality (\ref{S(n+2)}) is due to
Abott and Hanson \cite{AbbottHanson} and inequality (\ref{S(n+3)}) to Rowley \cite{RowleyRamsey}. 
The three following inequalities are our result.

\begin{align}
	S(n + 4) \geqslant &~111\,S(n) + 43 \label{S(n+4)}\\
	S(n + 5) \geqslant &~380\,S(n) + 148 \label{S(n+5)}\\
	S(n + 6) \geqslant &~1\,160\,S(n) + 536 \label{S(n+6)}
\end{align}

\par
Inequalities (\ref{S(n+1)}), (\ref{S(n+2)}) and (\ref{S(n+3)}) cannot be further improved (with this definition of S-template). 
Inequality (\ref{S(n+4)}) cannot be further improved by only searching for symmetric S-templates whose special color is the 
last in the order of apparition (and with a multiplicative factor less than or equal to 118). Inequality (\ref{S(n+5)}) can most 
likely be further improved but the improvement probably will not be substantial. Finally, a S-template corresponding to inequality 
(\ref{S(n+6)}) was found by extending into a S-template the Schur 6 partition used in \cite{rowley2021improved} (owing to the 
size of this template, its interest is limited and since it can easily be derived from above mentioned partition, it is not given in 
the appendix). Although we could not find a better S-template with seven colors, inequality (\ref{S(n+6)}) can definitely be 
improved by a wide margin. One may try to seek better templates by constraining less the search space and by using Monte-Carlo 
methods, as in \cite{Bouzy2015AnAP}. This could be the subject of a future work.

\par
The previous inequalities give new lower bounds for \(S(n)\) for
\( n \geqslant 9 \). We compute the lower
bounds for \( n \in [\![8,15]\!] \) using the four different inequalities. The best lower bounds are highlighted.

\renewcommand{\arraystretch}{0.2}

\begin{table}[H]

\label{LowerBoundsWS}
\[
\begin{array}{c}
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-2}
		\cellcolor{yellow}{3-3}
		\cellcolor{yellow}{4-4}
		\cellcolor{yellow}{4-5}
	\Body
		n & 8 & 9 & 10 & 11 \\
		33 \, S(n-3) + 6 & 5\,286 & 17\,694 & 55\,974 & 174\,444 \\
		111 \, S(n-4) + 43 & 4927 & 17\,803 & 59\,539 & 188\,299 \\
		380 \, S(n-5) + 148 & 5\,088 & 16\,868 & 60\,948 & 203\,828 \\
	\end{NiceArray}
	\\ \\
	\resetarraystretch
	\begin{NiceArray}{cwc{8ex}wc{10ex}wc{10ex}wc{11ex}}[hvlines]
	\CodeBefore
		\cellcolor{yellow}{2-3}
		\cellcolor{yellow}{4-2}
		\cellcolor{yellow}{4-4}
		\cellcolor{yellow}{4-5}
	\Body
		n & 12 & 13 & 14 & 15 \\
		33 \, S(n-3) + 6 & 587\,505 & 2\,011\,290 & 6\,726\,330 & 21\,272\,730 \\
		111 \, S(n-4) + 43 & 586\,789 & 1\,976\,176 & 6\,765\,271 & 22\,624\,951 \\
		380 \, S(n-5) + 148 & 644\,628 & 2\,008\,828 & 6\,765\,288 & 23\,160\,388 \\
	\end{NiceArray}
\end{array}
\]
\caption{New lower bounds for \( n \in [\![8,15]\!] \)}
\end{table}
\resetarraystretch

Except for \(S(8)\), \(S(9)\) and \(S(13)\), the best lower bounds are obtained thanks to
the third inequality \( S(n+5) \geqslant 380 \, S(n) + 148\). The table
doesn't go any further, but the same inequality allows to improve the
lower bounds for every \( n \geqslant 15 \).

\begin{corollary}
\begin{sloppypar}
The growth rate for Schur numbers (and Ramsey numbers \(R_n(3)\))  satisfies \({\gamma \geqslant \sqrt[5]{380} \approx 3.28}\).
\end{sloppypar}
\end{corollary}

\begin{proof}[\textsc{Proof.}]
It is a mere consequence of the inequality \( S(n+5) \geqslant 380 \, S(n) + 148\). As for Ramsey
numbers, the following inequality holds \(S(n) \leqslant R_n(3) - 2\) (see \cite{AbbottHanson}) hence the result. \\
\end{proof}


\subsection{Conclusion on S-templates}

In this section, we first formalized Rowley's template-based constructions \cite{RowleyRamsey} in the context of Schur numbers 
by introducing S-templates as well as a new sequence, \(S^+\). We provided relations between \(S^+\) and \(S\) then 
stated Rowley's construction method in the context of Schur numbers. We found new S-templates allowing us to obtain 
new lower bounds for schur numbers. One may notice that we mostly gave only lower bounds for \(S^+\). It should be possible to 
find better S-templates by making different assumptions or using a different method (Monte-Carlo methods for instance).

\par
In the next section, we provide similar results for weak Schur numbers. We introduce WS-templates and a corresponding sequence, 
\(\WS^+\). We then derive similar relations and a construction method allowing us to find new lower bounds for weak Schur numbers.