\documentclass[final,onefignum,onetabnum]{siamart190516}


\usepackage{amssymb}
\usepackage{dsfont}
\usepackage{MyMnSymbol}

\usepackage{float}

\usepackage{nicematrix}

\newsiamremark{remark}{Remark}

\newcommand{\WS}{\mathit{WS}}



\title{New lower bounds for Schur and weak Schur numbers \thanks{Submitted to the editors 02/24/2022.}}

\author{Romain Ageron\thanks{CentraleSup\'elec, Universit\'e{} Paris-Saclay, Gif-sur-Yvette, France
	(\email{romain.ageron@student-cs.fr}, \email{paul.casteras@student-cs.fr},
	\email{thibaut.pellerin@student-cs.fr}, \email{yann.portella@student-cs.fr}).}
\and Paul Casteras\footnotemark[2]
\and Thibaut Pellerin\footnotemark[2]
\and Yann Portella\footnotemark[2]
\and Arpad Rimmel\footnotemark[2]\ \thanks{LISN, CentraleSup\'elec, Universit\'e{} Paris-Saclay, Orsay, France
	(\email{arpad.rimmel@centralesupelec.fr}).}
\and Joanna Tomasik\footnotemark[2]\ \footnotemark[3]\ \thanks{Corresponding author
	(\email{joanna.tomasik@centralesupelec.fr}).}}

\headers{New lower bounds for Schur and weak Schur numbers}{R. Ageron, P. Casteras,
T. Pellerin, Y. Portella, A. Rimmel, J. Tomasik}


\begin{document}
\maketitle

\begin{abstract}
This article provides new lower bounds for both Schur and weak Schur numbers by exploiting a "template"-based approach. 
The concept of "template" is also generalized to weak Schur numbers. Finding new templates leads to explicit partitions 
improving lower bounds as well as the growth rate for Schur numbers, weak Schur numbers, and multicolor Ramsey numbers \(R_n(3)\). 
The new lower bounds include \(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS(6) \geqslant 646\), 
\(\WS (9) \geqslant 22\,536\) and \(\WS (10) \geqslant 71\,256 \).
\end{abstract}

\begin{keywords}
	Schur number, Weak Schur number, Ramsey theory, Sum-free partition
\end{keywords}

\begin{AMS}
	05D10, 11P81, 05A17, 11B75
\end{AMS}

\input{introduction.tex}

\input{schur.tex}

\input{weakschur.tex}

\section{Conclusion and future work}

We have produced new templates for Schur and Ramsey numbers. We also have generalized the concept of template to weak Schur numbers. 
These templates provide new general inequalities of the form \(S(n+k) \geqslant a S(n) + b\) and \(\WS(n+k) \geqslant a S(n) + b\) 
as well as new lower bounds for both Schur and weak Schur numbers. We have given a new lower bound \(\WS(6) \geqslant 646\) with 
a method that can yield slight improvements for fixed values of \(n\) over lower bounds obtained with the inequality 
\(\WS(n+1) \geqslant 4 S(n) + 2\).

Given that the introduction of templates is quite recent, we expect the bounds to be improved as this special type of partition 
becomes better understood and larger templates are found. One may try to produce better templates by using heuristics or 
approximation algorithms, such as Monte-Carlo algorithms for instance.

The best lower bound \(\WS(6) \geqslant 583\) achieved with a computer search using Monte-Carlo methods 
\cite{EliahouBook} is significantly lower than those obtained with a template-like structure: \(\WS(6) \geqslant 642\) \cite{RowleyWS}, 
\(\WS(6) \geqslant 646\). We have evidence that 583 is the maximal value in the search space considered by 
\cite{EliahouBook,Bouzy2015AnAP,Rafilipojaona}. Therefore, it questions the assumption that large partitions for \(\WS(n+1)\) 
start with large partitions for \(\WS(n)\). It also indicates that \(\WS(5)\) might need further investigation since the current lower 
bound \(\WS(5) \geqslant 196\) \cite{ELIAHOU2012175} was obtained by considering the same type of search space.


\section{Acknowledgments}

This study results from a long-term project performed within the framework of the Project Cluster "Training for Research" of CentraleSupélec. 
We would like to address warm thanks to Fred Rowley who kindly answered our many questions
and with who it was a pleasure to exchange. Last but not least, we would like to thank Marijn Heule for providing the backdoors
used in the computation of the new lower bound for \(\WS(6)\).


\bibliography{biblio}
\bibliographystyle{siamplain}

\include{appendix.tex}

\end{document}
